Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-07-27 doc/thesis/methods_results.Rnw
Defence version of thesis with changes

The attachment power parameter \gls{alpha} was varied among three values: 0.5,
1.0, and 1.5. For each value, the \software{sample\_pa} function in the
\software{igraph} package was used to simulate 100 networks, with the other
parameters set to \gls{N} = 5000 and \gls{m} = 2. This step yielded a total of
300 networks. An epidemic was simulated on each network using our
\software{nettree} binary until \gls{I} = 1000 nodes were infected, at which
point 500 of them were sampled to form a transmission tree. A total of 300
transmission trees were thus obtained, comprised of 100 trees for each of the
three values of \gls{alpha}. The trees were ``ladderized'' so that the subtree
descending from the left child of each node was not smaller than that
descending from the right child. Summary statistics, such as Sackin's index and
the ratio of internal to terminal branch lengths, were computed for each
simulated tree using our \software{treestat} binary. The trees were visualized
using the \software{ape} package~\autocite{paradis2004ape}. \add{Both the tree
kernel and the \gls{nltt} are pairwise statistics, and the \glspl{SVR}
classifiers we used to investigate them operate on pairwise distance matrices.}
Our \software{treekernel} binary was used to calculate the value of the kernel
for each pair of trees, with the meta-parameters set to $\lambda = 0.3$ and
$\sigma = 4$. These values were stored in a symmetric 300 $\times$ 300 kernel
matrix.  Similarly, we computed the \gls{nltt} statistic between each pair of
trees using our \software{treestat} binary, and stored them in a second $300
\times 300$ matrix.

To investigate the identifiability of \gls{alpha} from tree shape, we
constructed classifiers for \gls{alpha} based on the three tree shape
statistics discussed above. First, we used the \software{kernlab}
package~\autocite{zeileis2004kernlab} to create a \gls{kSVR} classifier using
the computed kernel matrix. Second, we used the \software{e1071}
package~\autocite{meyer2015e1071} to create an ordinary \gls{SVR} classifier
using the pairwise \gls{nltt} matrix. Finally, we performed an ordinary linear
regression of \gls{alpha} against Sackin's index. Each of these classifiers was
evaluated with 1000 two-fold cross-validations. We also performed a \gls{kPCA}
projection of the kernel matrix, and used it to visualize the separation of the
different \gls{alpha} values in the tree kernel's feature space. A schematic of
this experiment is presented in \cref{fig:kernelexpt}.

@@

\add{To assess the impact of the \gls{SMC} settings on \software{netabc}'s
accuracy, we ran \software{netabc} twice on the same simulated transmission
tree. For the first run, the \gls{SMC} settings were the same as in the other
simulations: 1000 particles, 5 simulated transmission trees per particle, and
\gls{alphaess} = 0.95. The second run was performed with 2000 particles, 10
simulated transmission trees per particle, and \gls{alphaess} = 0.99. To
investigate the extent to which errors in the estimated \gls{BA} parameters
were due to true features of the posterior, rather than an inaccurate \gls{ABC}
approximation, we performed marginal estimation for one set of parameter
values. Each combination of 1, 2, or 3 model parameters (14 combinations total)
was fixed to their known values, and the remaining parameters were estimated
with \software{netabc}. The parameter values were $\gls{alpha} = 0.0$, $\gls{m}
= 2$, $\gls{I} = 2000$, and $\gls{N} = 5000$.}

@@

    get.treestats <- function (param) {
        stats <- paste0("../../simulations/kernel-", param, "/statistics/*")
        d <- collect.data(stats)
        trees <- paste0("../../simulations/kernel-", param, "/tree/*")
        m <- collect.metadata(trees)
        rownames(m) <- sub("../../simulations/", "", rownames(m))
        d <- merge(d, m, by=0, suffixes=c("", ".1"))
        if (param == "alpha") {
            d <- subset(d, m == 2)
            colnames(d)[colnames(d) == "nsimnode"] <- "I"
        }
        setDT(d)
    }

    alpha.stats <- get.treestats("alpha")
    I.stats <- get.treestats("I")
    m.stats <- get.treestats("m")
    N.stats <- get.treestats("N")

@@

<<classifiers, include=FALSE>>=
    get.kernel.data <- function (param, step) {
        md <- collect.metadata(paste0('../../simulations/kernel-', param, '/', step, '/*'))
        if ('rbf_variance' %in% colnames(md)) {
          md <- subset(md, rbf_variance == 4 & decay_factor == 0.3)
        }
        if (param == 'alpha') {
            md <- subset(md, m == 2)
        }
        d <- setDT(collect.data(rownames(md)))
        if (param %in% c('alpha', 'm')) {
            d[,list(rsquared=mean(rsquared)), by=c('nsimnode', 'ntip')]
        } else if (param == 'I') {
            d[,list(rsquared=mean(rsquared)), by=ntip]
        } else {
            d[,list(rsquared=mean(rsquared)), by=c('I', 'ntip')]
        }
    }
    kernel.alpha <- get.kernel.data('alpha' ,'classifier')
    kernel.m <- get.kernel.data('m', 'classifier')
    kernel.I <- get.kernel.data('I', 'classifier')
    kernel.N <- get.kernel.data('N', 'classifier')

    sackin.alpha <- get.kernel.data('alpha', 'stats-classifier')
    sackin.m <- get.kernel.data('m', 'stats-classifier')
    sackin.I <- get.kernel.data('I', 'stats-classifier')
    sackin.N <- get.kernel.data('N', 'stats-classifier')

    nltt.alpha <- get.kernel.data('alpha', 'nltt-classifier')
    nltt.m <- get.kernel.data('m', 'nltt-classifier')
    nltt.I <- get.kernel.data('I', 'nltt-classifier')
    nltt.N <- get.kernel.data('N', 'nltt-classifier')
@

The accuracy of each classifier \add{on all \gls{BA} parameters is shown in}
\del{varied based on the parameter being tested} \cref{fig:rsquared}.
Classifiers based on \del{two other tree statistics,} the \gls{nltt} and
Sackin's index generally exhibited worse performance than the tree kernel,
although the magnitude of the disparity varied between the parameters
(\cref{fig:rsquared}, centre and right). The results were largely robust to
variations in the tree kernel meta-parameters \gls{lambda} and \gls{sigma},
although accuracy varied between different epidemic and sampling scenarios
(\cref{fig:alphacrossv,fig:mcrossv,fig:Icrossv,fig:Ncrossv}). \add{For all
parameters except \gls{m}, the absolute number of tips in the tree had a much
greater impact on accuracy than the proportion of infected individuals these
tips represented. However, for \gls{m}, both the number and proportion of
sampled tips had a strong impact or the accuracy of the \gls{kSVR}
(\cref{fig:mcrossv}).}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
    \caption[
        Cross-validation accuracy of kernel-SVR, nLTT-based SVR, and Sackin's
        index regression classifiers for BA model parameters.
    ]{
        Cross-validation accuracy of kernel-SVR classifier (left), SVR
        classifier using \gls{nltt} (centre), and linear regression using
        Sackin's index (right) for \gls{BA} model parameters. Kernel
        meta-parameters were set to $\gls{lambda} = 0.3$ and $\gls{sigma} = 4$.
        Each point was calculated based on 300 simulated transmission trees
        over networks with three different values of the parameter being
        tested, assuming perfect knowledge of the other parameters. Vertical
        lines are empirical 95\% confidence intervals based on 1000 two-fold
        cross-validations. \add{The classifiers for \gls{I} were not evaluated
        with 1000-tip trees, because one of the tested \gls{I} values was 500,
        and it is not possible to sample a tree of size 1000 from 500 infected
        individuals.}
    }
    \label{fig:rsquared}
\end{figure}

The \gls{kSVR} classifier for \gls{alpha} had an average $R^2$ of 
    \Sexpr{kernel.alpha[,round(mean(rsquared), 2)]},
compared to 
    \Sexpr{nltt.alpha[,round(mean(rsquared), 2)]}
for the \gls{nltt}-based SVR, and
    \Sexpr{sackin.alpha[,round(mean(rsquared), 2)]}
for the linear regression against Sackin's index. There was little variation
about the mean for different tree and epidemic sizes. No classifier could
accurately identify \gls{m} in any epidemic scenario, with average $R^2$ values
of 
  \Sexpr{kernel.m[,round(mean(rsquared), 2)]} for \gls{kSVR},
  \Sexpr{nltt.m[,round(mean(rsquared), 2)]} for the \gls{nltt}, and
  \Sexpr{sackin.m[,round(mean(rsquared), 2)]}
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the \gls{kSVR} was slightly higher
on 1000-tip trees 
    (average $R^2$ 
     \Sexpr{kernel.m[ntip == 100,round(mean(rsquared), 2)]},
     \Sexpr{kernel.m[ntip == 500,round(mean(rsquared), 2)]},
     \Sexpr{kernel.m[ntip == 1000,round(mean(rsquared), 2)]}
     for 100, 500, and 1000 tips respectively).

The accuracy of classifiers for \gls{I} \del{varied significantly
with}\add{appeared to be strongly influenced by} the number of tips in the
tree. For 100-tip trees, the average $R^2$ values were
  \Sexpr{kernel.I[ntip == 100, round(mean(rsquared), 2)]},
  \Sexpr{nltt.I[ntip == 100, round(mean(rsquared), 2)]}, and
  \Sexpr{sackin.I[ntip == 100, round(mean(rsquared), 2)]}
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  \Sexpr{kernel.I[ntip == 500, round(mean(rsquared), 2)]},
  \Sexpr{nltt.I[ntip == 500, round(mean(rsquared), 2)]}, and
  \Sexpr{sackin.I[ntip == 500, round(mean(rsquared), 2)]}.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVR} classifier ranged from
  \Sexpr{kernel.N[,round(min(rsquared), 2)]}
for the smallest epidemic and smallest sample size, to
  \Sexpr{kernel.N[,round(max(rsquared), 2)]}
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVR ranged from 
  \Sexpr{nltt.N[,round(min(rsquared), 2)]}
to
  \Sexpr{nltt.N[,round(max(rsquared), 2)]}.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  \Sexpr{sackin.N[,round(mean(rsquared), 2)]}
and little variation between scenarios.

@@

\del{The accuracy of grid search estimates largely paralleled that of the
\gls{kSVR} classifiers.} \Cref{fig:gridest} shows point estimates and
\add{50\%} and 95\% highest density intervals for each of the \gls{BA}
parameters, for one replicate experiment with 500-tip trees. Plots showing the
point estimates for all replicates can be found in
\cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all parameters
except \gls{m}, the error of point estimates was negatively correlated with the
number of sampled tips in the tree (for \gls{alpha}, \gls{I}, and \gls{N}
respectively: Spearman's $\rho$ = 
    \Sexpr{tip.cor["alpha", round(estimate, 2)]},
    \Sexpr{tip.cor["I", round(estimate, 2)]}, 
    \Sexpr{tip.cor["N", round(estimate, 2)]}; 
$p$-values
    $\Sexpr{tip.cor["alpha", pp(p.value)]}$,
    $\Sexpr{tip.cor["I", pp(p.value)]}$,
    $\Sexpr{tip.cor["N", pp(p.value)]}$).
The 95\% highest density intervals obtained for all parameters were extremely
wide, occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

@@

\add{Across all replicates, $R^2$ values for the correlations between the
estimated and true values were
    $\Sexpr{tests["alpha", round(rsquared, 2)]}$,
    $\Sexpr{tests["I", round(rsquared, 2)]}$,
    $\Sexpr{tests["m", round(rsquared, 2)]}$, and
    $\Sexpr{tests["N", round(rsquared, 2)]}$
for \gls{alpha}, \gls{I}, \gls{m}, and \gls{N} respectively. The mean absolute
errors of the point estimates were
    $\Sexpr{tests["alpha", round(abs.err, 2)]}$,
    $\Sexpr{tests["I", round(abs.err)]}$,
    $\Sexpr{tests["m", round(abs.err, 2)]}$, and
    $\Sexpr{tests["N", round(abs.err)]}$,
representing
    $\Sexpr{tests["alpha", round(grid.err)]}$\%,
    $\Sexpr{tests["I", round(grid.err)]}$\%,
    $\Sexpr{tests["m", round(grid.err)]}$\%, and
    $\Sexpr{tests["N", round(grid.err)]}$\%
of the respective grids. For \gls{I} and \gls{N}, relative errors may be more
appropriate to consider. An overestimate of 100 individuals would be very
misleading if the true population was only of size 100, but almost negligible
in a population of size 5000. The mean relative errors for \gls{I} and \gls{N}
respectively were
    $\Sexpr{tests["I", round(rel.err)]}$\% and 
    $\Sexpr{tests["N", round(rel.err)]}$\%.}

@@

\Cref{fig:abcptm2} shows \del{MAP} \add{posterior mean} point estimates of the
\gls{BA} model parameters obtained with \gls{ABC} on simulated data. The
estimates shown correspond only to the simulations where \gls{m} was set to 2,
however the results for $\gls{m} = 3$ and $\gls{m} = 4$ were similar
(\cref{fig:abcptm3,fig:abcptm4}). Average boundaries of 95\% \gls{HPD}
intervals are given in \cref{tab:abchpd}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{abc-point-estimate-m2}
    \caption[
        Posterior mean point estimates for \gls{BA} model parameters obtained
        by running \software{netabc} on simulated data, for simulations with
        $\gls{m} = 2$.
    ]{
        Posterior mean point estimates for \gls{BA} model parameters obtained
        by running \software{netabc} on simulated data, for simulations with
        $\gls{m} = 2$. Dashed lines indicate true values. (A) Estimates of
        \gls{alpha} and \gls{I} which were varied in these simulations against
        known values. (B) Estimates of \gls{m} and \gls{N} which were held
        fixed in these simulations at the values $\gls{m} = 2$ and 
        $\gls{N} = 5000$.
    }
    \label{fig:abcptm2}
\end{figure}

\begin{table*}[ht]
    \centering
    \input{\tablepath/abc-hpd.tex}
    \caption[
        Average posterior mean point estimates and 95\% \gls{HPD} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data.
    ]{
        Average posterior mean point estimates and 95\% \gls{HPD} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data. Three transmission trees were
        simulated under each combination of the listed parameter values, and
        the parameters were estimated with \gls{ABC} without training.
    }
    \label{tab:abchpd}
\end{table*}

@@

\add{To analyze the effects of the true parameter values on the
accuracy our estimates of \gls{alpha} and \gls{I}, we fitted one \gls{GLM}
for each of these two parameters, with error rate as the dependent variable and
the true parameter values as independent variables. Since the estimates of
\gls{m} and \gls{N} were roughly equal across all simulations
(\cref{fig:abcptm2,fig:abcptm3,fig:abcptm4}), \glspl{GLM} were not fitted for
these parameters. The estimated coefficients are shown in
\cref{tab:glmalpha,tab:glmI}. The \gls{GLM} were fitted using the inverse link
function. That is, if $p$ is the true value of the parameter and $\hat{p}$ is a
random variable representing our estimate of the parameter, the \gls{GLM}
posits a relationship of the form}
\[
    \madd{\E(|p-\hat{p}|) = (\beta_0 + \beta_\alpha \alpha + \beta_I I + \beta_m m)^{-1}},
\]
\add{where the $\beta$'s are coefficients to be fitted. If the true value
\gls{alpha}, say, is increased by one, the \emph{inverse} of the expected
absolute error will increase by $\beta_\alpha$. If $\beta_\alpha$ is positive,
it means that the absolute error decreases as the true value of \gls{alpha}
increases.}

@@

\add{The \gls{GLM} analysis indicated that the error in estimates of
\gls{alpha} decreased with larger true values of \gls{alpha} 
    ($p \Sexpr{pp(alpha.glm[Parameter == "alpha", p], eq=TRUE)}$)
and \gls{m}
    ($p \Sexpr{pp(alpha.glm[Parameter == "m", p], eq=TRUE)}$)
but was not significantly affected by \gls{I} (\cref{tab:glmalpha}).
Qualitatively, \gls{alpha} seemed to be only weakly identifiable between the
values of 0 and 0.5 (\cref{fig:abcptm2}). The error in the estimated prevalence
\gls{I} was slightly lower for smaller values of \gls{alpha}
    ($p \Sexpr{pp(I.glm[Parameter == "alpha", p], eq=TRUE)}$)
and \gls{I}
    ($p \Sexpr{pp(I.glm[Parameter == "I", p], eq=TRUE)}$),
but was not significantly affected by the true value of \gls{m}
(\cref{tab:glmI}).}

\del{The accuracy of the parameter estimates obtained with ABC paralleled the
results from the \gls{kSVR} classifier. Of the four parameters, $\alpha$ was
the most accurately estimated, with point estimates having a median [IQR]
absolute error of 
    \Sexpr{d[,round(median(alpha_error), 2)]} 
    [\Sexpr{d[,round(quantile(alpha_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(alpha_error, 0.75), 2)]}].
The errors when the true value of $\alpha$ was zero were significantly greater
than those for the other values 
    (Wilcoxon rank-sum test, $p$ = 
    $\Sexpr{latexSN(round(zero.alpha.test[["p.value"]], 4))}$).
Errors in estimating $\alpha$ did not vary across the true values of $m$ or $I$
(both one-way ANOVA).}

>>>>

The attachment power parameter \gls{alpha} was varied among three values: 0.5,
1.0, and 1.5. For each value, the \software{sample\_pa} function in the
\software{igraph} package was used to simulate 100 networks, with the other
parameters set to \gls{N} = 5000 and \gls{m} = 2. This step yielded a total of
300 networks. An epidemic was simulated on each network using our
\software{nettree} binary until \gls{I} = 1000 nodes were infected, at which
point 500 of them were sampled to form a transmission tree. A total of 300
transmission trees were thus obtained, comprised of 100 trees for each of the
three values of \gls{alpha}. The trees were ``ladderized'' so that the subtree
descending from the left child of each node was not smaller than that
descending from the right child. Summary statistics, such as Sackin's index and
the ratio of internal to terminal branch lengths, were computed for each
simulated tree using our \software{treestat} binary. The trees were visualized
using the \software{ape} package~\autocite{paradis2004ape}. \add{Both the tree
kernel and the \gls{nltt} are pairwise statistics, and the \glspl{SVM}
classifiers we used to investigate them operate on pairwise distance matrices.}
Our \software{treekernel} binary was used to calculate the value of the kernel
for each pair of trees, with the meta-parameters set to $\lambda = 0.3$ and
$\sigma = 4$. These values were stored in a symmetric 300 $\times$ 300 kernel
matrix.  Similarly, we computed the \gls{nltt} statistic between each pair of
trees using our \software{treestat} binary, and stored them in a second $300
\times 300$ matrix.

To investigate the identifiability of \gls{alpha} from tree shape, we
constructed classifiers for \gls{alpha} based on the three tree shape
statistics discussed above. The \software{kernlab}
package~\autocite{zeileis2004kernlab} was used to create a \gls{kSVM}
classifier using the computed kernel matrix, and the \software{e1071}
package~\autocite{meyer2015e1071} was used to create ordinary \gls{SVM}
classifiers using the pairwise \gls{nltt} matrix and Sackin's index values.
\del{Finally, we performed an ordinary linear regression of \gls{alpha} against
Sackin's index.} Each of these classifiers was evaluated with 1000 two-fold
cross-validations \add{with equally-sized folds}. We also performed a
\gls{kPCA} projection of the kernel matrix, and used it to visualize the
separation of the different \gls{alpha} values in the tree kernel's feature
space. A schematic of this experiment is presented in \cref{fig:kernelexpt}.

@@

\add{To assess the impact of the \gls{SMC} settings on \software{netabc}'s
accuracy, we ran \software{netabc} twice on the same simulated transmission
tree. For the first run, the \gls{SMC} settings were the same as in the other
simulations: 1000 particles, 5 simulated transmission trees per particle, and
\gls{alphaess} = 0.95. The second run was performed with 2000 particles, 10
simulated transmission trees per particle, and \gls{alphaess} = 0.97. To
investigate the extent to which errors in the estimated \gls{BA} parameters
were due to true features of the posterior, rather than an inaccurate \gls{ABC}
approximation, we performed marginal estimation for one set of parameter
values. Each combination of 1, 2, or 3 model parameters (14 combinations total)
was fixed to their known values, and the remaining parameters were estimated
with \software{netabc}. The parameter values were $\gls{alpha} = 0.0$, $\gls{m}
= 2$, $\gls{I} = 2000$, and $\gls{N} = 5000$.}

@@

    get.treestats <- function (param) {
        stats <- paste0("../../simulations/kernel-", param, "/statistics/*")
        d <- collect.data(stats)
        trees <- paste0("../../simulations/kernel-", param, "/tree/*")
        m <- collect.metadata(trees)
        rownames(m) <- sub("../../simulations/", "", rownames(m))
        rownames(m) <- sub(".bk.2", "", rownames(m))
        d <- merge(d, m, by=0, suffixes=c("", ".1"))
        if (param == "alpha") {
            d <- subset(d, m == 2)
            colnames(d)[colnames(d) == "nsimnode"] <- "I"
        }
        setDT(d)
    }

    alpha.stats <- get.treestats("alpha")
    I.stats <- get.treestats("I")
    m.stats <- get.treestats("m")
    N.stats <- get.treestats("N.bk.2")

@@

<<classifiers, include=FALSE>>=
    get.kernel.data <- function (param, step) {
        md <- collect.metadata(paste0('../../simulations/kernel-', param, '/', step, '/*'))
        if ('rbf_variance' %in% colnames(md)) {
          md <- subset(md, rbf_variance == 4 & decay_factor == 0.3)
        }
        if (param == 'alpha') {
            md <- subset(md, m == 2)
        }
        d <- setDT(collect.data(rownames(md)))
        if (param %in% c('alpha', 'm')) {
            d[,list(rsquared=mean(rsquared), accuracy=mean(accuracy)), by=c('nsimnode', 'ntip')]
        } else if (param == 'I') {
            d[,list(rsquared=mean(rsquared), accuracy=mean(accuracy)), by=ntip]
        } else {
            d[,list(rsquared=mean(rsquared), accuracy=mean(accuracy)), by=c('I', 'ntip')]
        }
    }
    kernel.alpha <- get.kernel.data('alpha' ,'classifier')
    kernel.m <- get.kernel.data('m', 'classifier')
    kernel.I <- get.kernel.data('I', 'classifier')
    kernel.N <- get.kernel.data('N.bk.2', 'classifier')

    sackin.alpha <- get.kernel.data('alpha', 'stats-classifier')
    sackin.m <- get.kernel.data('m', 'stats-classifier')
    sackin.I <- get.kernel.data('I', 'stats-classifier')
    sackin.N <- get.kernel.data('N.bk.2', 'stats-classifier')

    nltt.alpha <- get.kernel.data('alpha', 'nltt-classifier')
    nltt.m <- get.kernel.data('m', 'nltt-classifier')
    nltt.I <- get.kernel.data('I', 'nltt-classifier')
    nltt.N <- get.kernel.data('N.bk.2', 'nltt-classifier')
@

The accuracy of each classifier, which is the proportion of trees assigned the
correct parameter value, \add{is shown in} \del{varied based on the parameter
being tested} \cref{fig:rsquared} \add{Since there were three possible values,
random guessing would produce an accuracy of 0.33.} Classifiers based on
\del{two other tree statistics,} the \gls{nltt} and Sackin's index generally
exhibited worse performance than the tree kernel, although the magnitude of the
disparity varied between the parameters (\cref{fig:rsquared}, centre and
right). The results were largely robust to variations in the tree kernel
meta-parameters \gls{lambda} and \gls{sigma}, although accuracy varied between
different epidemic and sampling scenarios
(\cref{fig:alphacrossv,fig:mcrossv,fig:Icrossv,fig:Ncrossv}). \add{For all
parameters except \gls{m}, the absolute number of tips in the tree had a much
greater impact on accuracy than the proportion of infected individuals these
tips represented. However, for \gls{m}, both the number and proportion of
sampled tips had a strong impact or the accuracy of the \gls{kSVM}
(\cref{fig:mcrossv}).}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
    \caption[
        Cross-validation accuracy of kernel-SVM, nLTT-based SVM, and Sackin's
        SVM classifiers for BA model parameters.
    ]{
        Cross-validation accuracy of kernel-SVM classifier (left), SVM
        classifier using \gls{nltt} (centre), and SVM using Sackin's index
        (right) for \gls{BA} model parameters. Kernel meta-parameters were set
        to $\gls{lambda} = 0.3$ and $\gls{sigma} = 4$. Each point was
        calculated based on 300 simulated transmission trees over networks with
        three different values of the parameter being tested, assuming perfect
        knowledge of the other parameters. Vertical lines are empirical 95\%
        confidence intervals based on 1000 two-fold cross-validations. \add{The
        classifiers for \gls{I} were not evaluated with 1000-tip trees, because
        one of the tested \gls{I} values was 500, and it is not possible to
        sample a tree of size 1000 from 500 infected individuals.}
    }
    \label{fig:rsquared}
\end{figure}

The \gls{kSVM} classifier for \gls{alpha} had an average accuracy of 
    \Sexpr{kernel.alpha[,round(mean(accuracy), 2)]},
compared to 
    \Sexpr{nltt.alpha[,round(mean(accuracy), 2)]}
for the \gls{nltt}-based SVM, and
    \Sexpr{sackin.alpha[,round(mean(accuracy), 2)]}
for Sackin's index. There was little variation about the mean for different
tree and epidemic sizes. No classifier could accurately identify \gls{m} in any
epidemic scenario, with average accuracy values of 
  \Sexpr{kernel.m[,round(mean(accuracy), 2)]} for \gls{kSVM},
  \Sexpr{nltt.m[,round(mean(accuracy), 2)]} for the \gls{nltt}, and
  \Sexpr{sackin.m[,round(mean(accuracy), 2)]}
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the \gls{kSVM} was slightly higher
on 1000-tip trees 
    (average 
     \Sexpr{kernel.m[ntip == 100,round(mean(accuracy), 2)]},
     \Sexpr{kernel.m[ntip == 500,round(mean(accuracy), 2)]},
     \Sexpr{kernel.m[ntip == 1000,round(mean(accuracy), 2)]}
     for 100, 500, and 1000 tips respectively).

The accuracy of classifiers for \gls{I} \del{varied significantly
with}\add{appeared to be strongly influenced by} the number of tips in the
tree. For 100-tip trees, the average accuracy values were
  \Sexpr{kernel.I[ntip == 100, round(mean(accuracy), 2)]},
  \Sexpr{nltt.I[ntip == 100, round(mean(accuracy), 2)]}, and
  \Sexpr{sackin.I[ntip == 100, round(mean(accuracy), 2)]}
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  \Sexpr{kernel.I[ntip == 500, round(mean(accuracy), 2)]},
  \Sexpr{nltt.I[ntip == 500, round(mean(accuracy), 2)]}, and
  \Sexpr{sackin.I[ntip == 500, round(mean(accuracy), 2)]}.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The accuracy of the \gls{kSVM} classifier ranged from
  \Sexpr{kernel.N[,round(min(accuracy), 2)]}
for the smallest epidemic and smallest sample size, to
  \Sexpr{kernel.N[,round(max(accuracy), 2)]}
for the largest. Likewise, accuracy for the \gls{nltt}-based SVM ranged from 
  \Sexpr{nltt.N[,round(min(accuracy), 2)]}
to
  \Sexpr{nltt.N[,round(max(accuracy), 2)]}.
Sackin's index did not accurately classify $N$ in any scenario, with an average
accuracy of
  \Sexpr{sackin.N[,round(mean(accuracy), 2)]}
and little variation between scenarios.

@@

\del{The accuracy of grid search estimates largely paralleled that of the
\gls{kSVM} classifiers.} \Cref{fig:gridest} shows point estimates and
\add{50\%} and 95\% highest density intervals for each of the \gls{BA}
parameters, for one replicate experiment with 500-tip trees. Plots showing the
point estimates for all replicates can be found in
\cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all parameters
except \gls{m}, the error of point estimates was negatively correlated with the
number of sampled tips in the tree (for \gls{alpha}, \gls{I}, and \gls{N}
respectively: Spearman's $\rho$ = 
    \Sexpr{tip.cor["alpha", round(estimate, 2)]},
    \Sexpr{tip.cor["I", round(estimate, 2)]}, 
    \Sexpr{tip.cor["N", round(estimate, 2)]}; 
$p$-values
    $\Sexpr{tip.cor["alpha", pp(p.value)]}$,
    $\Sexpr{tip.cor["I", pp(p.value)]}$,
    $\Sexpr{tip.cor["N", pp(p.value)]}$).
The 95\% highest density intervals obtained for all parameters were extremely
wide, occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

@@

\add{Across all replicates, $R^2$ values for the correlations between the
estimated and true values were
    $\Sexpr{tests["alpha", round(rsquared, 2)]}$,
    $\Sexpr{tests["I", round(rsquared, 2)]}$,
    $\Sexpr{tests["m", round(rsquared, 2)]}$, and
    $\Sexpr{tests["N", round(rsquared, 2)]}$
for \gls{alpha}, \gls{I}, \gls{m}, and \gls{N} respectively. The mean absolute
errors of the point estimates were
    $\Sexpr{tests["alpha", round(abs.err, 2)]}$,
    $\Sexpr{tests["I", round(abs.err)]}$,
    $\Sexpr{tests["m", round(abs.err, 2)]}$, and
    $\Sexpr{tests["N", round(abs.err)]}$,
representing
    $\Sexpr{tests["alpha", round(grid.err)]}$\%,
    $\Sexpr{tests["I", round(grid.err)]}$\%,
    $\Sexpr{tests["m", round(grid.err)]}$\%, and
    $\Sexpr{tests["N", round(grid.err)]}$\%
of the respective grids. Note that \cref{fig:gridest} contains only one
replicate data point per parameter value, out of ten total. For \gls{I} and
\gls{N}, relative errors may be more appropriate to consider. An overestimate
of 100 individuals would be very misleading if the true population was only of
size 100, but almost negligible in a population of size 5000. The mean relative
errors for \gls{I} and \gls{N} respectively were
    $\Sexpr{tests["I", round(rel.err)]}$\% and 
    $\Sexpr{tests["N", round(rel.err)]}$\%.}

@@

\Cref{fig:abcpt} shows \del{MAP} \add{stratified posterior mean} point
estimates of the \gls{BA} model parameters \add{\gls{alpha} and \gls{I}}
obtained with \gls{ABC} on simulated data. \add{The parameters $m$ and $N$ were
not identifiable with ABC for any parameter combinations (\cref{fig:abcpt2}).}
\del{The estimates shown correspond only to the simulations where \gls{m} was
set to 2, however the results for $\gls{m} = 3$ and $\gls{m} = 4$ were
similar.} Average boundaries of 95\% \gls{HPD} intervals are given in
\cref{tab:abchpd}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{abc-boxplot}
    \caption[
        Posterior mean point estimates for BA model parameters $\alpha$ and $I$
        obtained by running \textit{netabc} on simulated data, stratified by
        true parameter values.
    ]{
        Posterior mean point estimates for BA model parameters $\alpha$ and $I$
        obtained by running \textit{netabc} on simulated data, stratified by
        true parameter values. First row of plots contains true versus
        estimated values of $\alpha$; second row contains true versus estimated
        values of $I$. Columns are stratified by $\alpha$, $I$, and $m$
        respectively. Dashed lines indicate true values. 
    }
    \label{fig:abcpt}
\end{figure}

\begin{table}[ht]
    \centering
    \input{\tablepath/abc-hpd.tex}
    \caption[
        Average posterior mean point estimates and 95\% \gls{HPD} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data.
    ]{
        Average posterior mean point estimates and 95\% \gls{HPD} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data. Three transmission trees were
        simulated under each combination of the listed parameter values, and
        the parameters were estimated with \gls{ABC} without training.
    }
    \label{tab:abchpd}
\end{table}

@@

\add{To analyze the effects of the true parameter values on the
accuracy our estimates of \gls{alpha} and \gls{I}, we fitted one \gls{GLM}
for each of these two parameters, with error rate as the dependent variable and
the true parameter values as independent variables. Since the estimates of
\gls{m} and \gls{N} were roughly equal across all simulations
(\cref{fig:abcpt2}), \glspl{GLM} were not fitted for these parameters. The
estimated coefficients are shown in \cref{tab:glmalpha,tab:glmI}. The \gls{GLM}
were fitted using the inverse link function. That is, if $p$ is the true value
of the parameter and $\hat{p}$ is a random variable representing our estimate
of the parameter, the \gls{GLM} posits a relationship of the form}
\[
    \madd{\E(|p-\hat{p}|) = (\beta_0 + \beta_\alpha \alpha + \beta_I I + \beta_m m)^{-1}},
\]
\add{where the $\beta$'s are coefficients to be fitted. If the true value
\gls{alpha}, say, is increased by one, the \emph{inverse} of the expected
absolute error will increase by $\beta_\alpha$. If $\beta_\alpha$ is positive,
it means that the absolute error decreases as the true value of \gls{alpha}
increases.}

@@

\add{The \gls{GLM} analysis indicated that the error in estimates of
\gls{alpha} decreased with larger true values of \gls{alpha} 
    ($p \Sexpr{pp(alpha.glm[Parameter == "alpha", p], eq=TRUE)}$)
and \gls{m}
    ($p \Sexpr{pp(alpha.glm[Parameter == "m", p], eq=TRUE)}$)
but was not significantly affected by \gls{I} (\cref{tab:glmalpha}).
Qualitatively, \gls{alpha} seemed to be only weakly identifiable between the
values of 0 and 0.5 (\cref{fig:abcpt}). The error in the estimated prevalence
\gls{I} was slightly lower for smaller values of \gls{alpha}
    ($p \Sexpr{pp(I.glm[Parameter == "alpha", p], eq=TRUE)}$)
and \gls{I}
    ($p \Sexpr{pp(I.glm[Parameter == "I", p], eq=TRUE)}$),
but was not significantly affected by the true value of \gls{m}
(\cref{tab:glmI}).}

\del{The accuracy of the parameter estimates obtained with ABC paralleled the
results from the \gls{kSVM} classifier. Of the four parameters, $\alpha$ was
the most accurately estimated, with point estimates having a median [IQR]
absolute error of 
    \Sexpr{d[,round(median(alpha_error), 2)]} 
    [\Sexpr{d[,round(quantile(alpha_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(alpha_error, 0.75), 2)]}].
The errors when the true value of $\alpha$ was zero were significantly greater
than those for the other values 
    (Wilcoxon rank-sum test, $p$ = 
    $\Sexpr{latexSN(round(zero.alpha.test[["p.value"]], 4))}$).
Errors in estimating $\alpha$ did not vary across the true values of $m$ or $I$
(both one-way ANOVA).}
